---
output: reprex::reprex_document
knit: reprex::reprex_render
---

Download the necessarty packages and set the working directory to where the data is stored.

```{r}
install.packages("LexisNexisTools")
install.packages("quanteda.textmodels")
install.packages("BERTopic")
install.packages("reticulate")
```

```{r}
library("LexisNexisTools")
setwd("D:/JADS/Thesis/Data")
```
 
Rename files using info from cover pages

```{r}
report <- lnt_rename(x = getwd(), report = TRUE, recursive = TRUE)
```

Read all Lexis files in this folder

```{r}
LNToutput <- lnt_read(x = getwd(), recursive = TRUE)
```

Extract data frames

```{r}
meta_df       <- LNToutput@meta        # one row per article with metadata
articles_df   <- LNToutput@articles    # article-level text
paragraphs_df <- LNToutput@paragraphs  # paragraph-level text
```

## Writing the data to csv files for later use

```{r}
# Load libraries first
library(dplyr)
library(quanteda)
library(readr)

# Define stopwords once
my_stop <- c(
  stopwords("nl"),
  stopwords("en"),
  "ık", "ın", "zıjn", "ıs", "wıe", "ıets", "ıemand",
  "zıch", "dıt", "pagina", "we", "waar", "jaar", "wel"
)

# Reusable cleaner
remove_stopwords_text <- function(x, sw) {
  toks_local <- tokens(
    x,
    remove_punct = TRUE,
    remove_symbols = TRUE,
    remove_separators = TRUE
  ) %>%
    tokens_remove(sw) %>%
    tokens_remove(pattern = "^[[:punct:]]+$") %>%
    tokens_remove(pattern = "^[a-z]$")
  
  sapply(toks_local, paste, collapse = " ")
}

# Build corpus_df from all metadata + article text only
corpus_df <- meta_df %>%
  left_join(
    articles_df %>% dplyr::select(ID, Article),
    by = "ID"
  ) %>%
  dplyr::filter(!is.na(Newspaper), !is.na(Date), !is.na(Article)) %>%
  mutate(Article = remove_stopwords_text(as.character(Article), my_stop))

# Optional: save directly from this canonical block
readr::write_csv(corpus_df, "D:/JADS/Thesis/corpus_df_processed.csv")

# Create quanteda corpus
corp <- quanteda::corpus(corpus_df, text_field = "Article")
quanteda::docvars(corp) <- corpus_df %>% dplyr::select(-Article)
quanteda::docnames(corp) <- corpus_df$ID
```

Analyze the Big Data

```{r}
library(dplyr)
library(lubridate)
library(stringr)
library(ggplot2)
library(tibble)

year_target <- 2024  # change to 2017 (or any year) when needed

# 1) Quick diagnostic
cat("Date column class:", class(meta_df$Date), "\n")
print(head(meta_df$Date, 10))

# 2) Robust date parsing
date_std <- if (inherits(meta_df$Date, "Date")) {
  as.Date(meta_df$Date)
} else {
  raw_date <- as.character(meta_df$Date) %>%
    str_squish() %>%
    str_replace_all(",", "")  # remove all commas

  suppressWarnings(
    as.Date(
      parse_date_time(
        raw_date,
        orders = c(
          "Y-m-d", "d-m-Y", "d/m/Y", "m/d/Y", "Ymd",
          "d b Y", "d B Y", "b d Y", "B d Y", "d-b-Y", "d-B-Y",
          "Y-m-d H:M:S", "d/m/Y H:M:S", "m/d/Y H:M:S"
        ),
        locale = "C"
      )
    )
  )
}

# 3) Parsing quality
parse_rate <- round(mean(!is.na(date_std)) * 100, 2)
cat("Parsed dates:", sum(!is.na(date_std)), "of", length(date_std), "(", parse_rate, "%)\n")

# 4) Month popularity for selected year
popular_months <- tibble(Date = date_std) %>%
  filter(!is.na(Date), year(Date) == year_target) %>%
  mutate(month_num = month(Date)) %>%
  count(month_num, name = "n") %>%
  mutate(
    month = month(month_num, label = TRUE, abbr = FALSE),
    pct = round(100 * n / sum(n), 2)
  ) %>%
  arrange(desc(n), month_num)

if (nrow(popular_months) == 0) {
  cat("No parsed dates found for year", year_target, "\n")
} else {
  print(popular_months)

  # 5) Most popular month(s), including ties
  top_months <- popular_months %>%
    filter(n == max(n))

  cat("\nMost popular month(s) in", year_target, ":\n")
  print(top_months)

  # 6) Bar chart in calendar order
  plot_df <- popular_months %>%
    arrange(month_num) %>%
    mutate(month = factor(as.character(month), levels = as.character(month)))

  ggplot(plot_df, aes(x = month, y = n)) +
    geom_col(fill = "#2C7FB8") +
    labs(
      title = paste("Article count by month (", year_target, ")", sep = ""),
      x = "Month",
      y = "Number of articles"
    ) +
    theme_minimal(base_size = 12)
}
```

Pre-proccessing for NLP analysis

```{r}
# Load libraries first
library(dplyr)
library(quanteda)

# Create corpus_df by combining metadata with article text
corpus_df <- meta_df %>%
  left_join(articles_df, by = c("ID" = "ID"))

# Filter for missing values
corpus_df <- corpus_df %>%
  dplyr::filter(!is.na(Newspaper), !is.na(Date))

#set.seed(123)  # For reproducibility
#corpus_df <- corpus_df %>%
#  slice_sample(n = 2000)

# Create quanteda corpus
corp <- quanteda::corpus(corpus_df, text_field = "Article")

# Add document variables (metadata)
quanteda::docvars(corp) <- corpus_df %>% 
  dplyr::select(-Article)

# Set document names
quanteda::docnames(corp) <- corpus_df$ID
```

Tokenization

```{r}
docnames(corp) <- rownames(corpus_df)

my_stop <- c(
  stopwords("nl"),
  stopwords("en"),
  "ık", "ın", "zıjn", "ıs", "wıe", "ıets", "ıemand",
  "zıch", "dıt", "pagina", "we", "waar", "jaar", "wel"
)

# Tokenize with punctuation/symbol removal
toks <- tokens(corp,
               remove_punct   = TRUE,
               remove_symbols = TRUE,
               remove_separators = TRUE) %>%
  tokens_remove(my_stop) %>%
  tokens_remove(pattern = "^[[:punct:]]+$") %>%
  tokens_remove(pattern = "^[a-z]$")  # Remove single letters
```

Save the processed corpus for later use in topic modelling and supervised models.  

```{r}
readr::write_csv(corpus_df, "D:/JADS/Thesis/corpus_df_processed.csv")
``` 

Then, we create a document-feature matrix (DFM) -- a base for topic models and supervised models.

```{r}
dfmat <- dfm(toks) %>%
  dfm_trim(min_termfreq = 10, min_docfreq = 5)
``` 

## Structured topic modelling

We use topic modelling to discover themes.

```{r}
library(stm)

stm_in   <- quanteda::convert(dfmat, to = "stm")
stm_meta <- quanteda::docvars(corp)[quanteda::docnames(dfmat), , drop = FALSE] %>%
  as.data.frame()

# keep date numeric for s(Date)
stm_meta$Date <- as.numeric(as.Date(stm_meta$Date))

stopifnot(length(stm_in$documents) == nrow(stm_meta))

mod <- stm::stm(
  documents  = stm_in$documents,
  vocab      = stm_in$vocab,
  K          = 20,
  prevalence = ~ Newspaper * s(Date),
  data       = stm_meta,
  init.type  = "Spectral"
)
```

Inspect the top words in the first 5 topics

```{r}
labelTopics(mod, topics = 1:5)
````

Topic distribution by newspaper
